Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalog Data

### 2012-10-31 doc/paper-lcr.tex
abstract done - shipping to collaborators!

\todo{PJM}{Make abstract match conclusions.}

Strong gravitational lenses can provide high precision cosmological distance
measurements; one of the most important sources of systematic error in these
estimates is the additional mass along the line of sight to the source. Weak
lensing perturbations by galaxies and groups in this beam can contribute an ``external
convergence'' $\kappa$ that dominates the  uncertainty in the inferred
distances; likewise, it may also contaminate estimates of the luminosity of
objects at high redshift (such as reionisation-era proto-galaxies, and type Ia
supernovae standard candles).  We characterise this uncertainty by marginalising
over a probability density function for $\kappa$: here, we investigate the use
of a simple halo model to estimate $\Pr(\kappa)$ given noisy estimates of the
photometric redshift and stellar mass of galaxies observed in optical imaging
surveys of the field in question. We use mock catalogs from the Millenium
Simulation and compare our reconstructed $\Pr(\kappa)$ to ray-traced $\kappa$ values,
we find that our model produces a precise estimator of $\kappa$ with only small biases.
For each lightcone the individual $\Pr(\kappa)$ from this reconstruction process are
typically 1.8 times more precise than the global distribution of $\kappa$ given perfect knowledge 
of the halo mass and redshift. After incorporating uncertainties on redshift and halo and stellar masses
we typically find that the reconstruction yields an improvement factor of 1.4. For each line of sight the
 $\Pr(\kappa)$ can be formed before the investment of light-curve monitoring: by following up only the 
most constrained lines of sight a factor of 2 improvement over the global $\pr(\kappax)$ should be 
achievable.

>>>>

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappa$ that
in some cases dominates the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappa$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappa|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogs from the Millennium Simulation, and compare
our reconstructed $\pr(\kappa|\data)$ to ray-traced $\kappa$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields an improvement in precision of
40\% over the global $\pr(\kappa)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations: we find that selecting systems by the precision with which
$\kappa$ can be estimated leads to a low density sample that would give
distance estimates a factor of 2 more precise, with biases as low as 0.5\%.
Preferentially selecting lines of sight with high external shear (as might
happen in an assembled sample of quadruply-imaged time delay lenses) has the
opposite effect, with distances biased by around 1\%. We suggest areas for the
improvement of this general analysis framework that should allow yet more
accurate cosmological inferences to be made.
